---
title: "Data Visualization with R using ggplot2"
author: "Lukman Aliyu Jibril"
date: "2023-10-07"
categories: [data science, r programming,data visualization]
---
In this article, I attempt to briefly display the ease of data visualization using the ggplot2 library in R. 
As always, the first thing is to import the library using the `library` syntax. Here, I import tidyverse which includes ggplot2 as a core member of the tidyverse. I also import `palmerpenguins` as it contains the penguins dataset I will like to use. 

```{r}
library(tidyverse)
library(palmerpenguins)
```

Looking closely at the `mpg` dataset to see what variables it contains. 

```{r}
glimpse(mpg)
```

Checking out the `palmer penguins` dataset too. 

```{r}
glimpse(penguins)
```

Now, I use the easy to follow ggplot syntax to plot mpg dataset, `displ` against `hwy`. I also introduce a smooth line to accurately see the way the variables relate. 
```{r}
ggplot(data=mpg,mapping=aes(x= hwy, y= displ)) + geom_point() + geom_smooth(method= "lm")
```

Second plot is on the penguins data. I compare the bill length and bill depth and see how species affect the relationship. 
```{r}
ggplot(
  data = penguins,
  mapping = aes(
    x = bill_length_mm, y = bill_depth_mm, 
    color = species, shape = species
  )
) +
  geom_point() +
  labs(color = "Species") +
  guides(
    shape = guide_legend(title = "Species", override.aes = list(color = NULL))
  )

```


## Conclusion

The ggplot2 library offers an easy way to do data visualization in R. However, it is important to note that mastering visualization requires a lot of practice and a single article is not enough. Let's all keep learning friends.