---
title: "The Case for Writing Vectorised Code in R"
author: "Lukman Aliyu Jibril"
date: "2023-07-29"
categories: [r programming, vectorization]
---

## Writing Vectorized Code
The first thing is to define vectorized code. According to Garrett Grolemund in his book "Hands-On Programming with R", vectorised code has the following qualities:

- Logical tests
- Subsetting
- Element-wise execution

The key thing about vectorization is that operations are performed on all elements of a vector simultaneously without the need of explicitly looping over the elements of the vector. This can make R code significantly faster. 

As an illustration, let's look at two functions that do the same thing. One is vectorized and other is not. We will then compare the time it takes for both functions to be executed to determine which technique is more efficient. 

###  Regular Implementation Using a For Loop (not vectorized)

```{r}
abs_loop <- function(vec){
  for (i in 1:length(vec)) {
    if (vec[i] < 0) {
      vec[i] <- -vec[i]
    }
  }
  vec
}
```


###  Vectorized Implementation

```{r}
abs_sets <- function(vec){
  negs <- vec < 0
  vec[negs] <- vec[negs] * -1
  vec
}
```

Having defined the two functions, the next thing is to define the variable that will be passed to the both functions, as shown below:

```{r}
long <- rep(c(-1, 1), 5000000)

```


Finally, the time it takes for each function to run is calculated and compared: 


# Unvectorized 
```{r}
system.time(abs_loop(long))

```


# Vectorized
```{r}
system.time(abs_sets(long))

```

It can be seen that the vectorized function (ab_sets()) is about three times faster. The speed advantage increases as the size of the variable (long) increases. 
Imagine working on large vectors and the time that will be saved by vectorizing all our computations. Even the computational costs will be significantly reduced. 

In conclusion, since R is built around vectors and most R functions are vectorized, using vectorization will be a significant addition to the toolkit of every R user and this can significantly improve productivity. 